---
title: "Homework 6"
author: "Eileen Shea"
date: "November 27, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(modelr)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1

First we need to read in the data.

```{r}
homicide_data = read_csv("./data/homicide-data.csv")
```

Next we will create a `city_state` variable and a binary variable indicating whether the homicide is solved. We will also omit the cities Dallas, TX; Phoenix, AZ; Kansas City, MO; and Tulsa, AL. Lastly, we will modify `victim_race` to have categories white and non-white, with white as the reference category, and change victim_age to a numeric variable.

```{r}
US_hom_data = homicide_data %>% 
  mutate(state = toupper(state), city_state = str_c(city, state, sep = ", "),
        resolved = ifelse(disposition == "Closed without arrest" | disposition == "Open/No arrest", FALSE, TRUE)) %>% 
  filter(city_state != "Dallas, TX", city_state != "Phoenix, AZ", city_state != "Kansas City, MO", city_state != "Tulsa, AL") %>% 
  mutate(victim_race = ifelse(victim_race == "White", "White", "Non-White"),
         victim_race = factor(victim_race, levels = c("White", "Non-White")),
         victim_age  = as.numeric(victim_age))
```

Now for the city of Baltimore, MD we will fit a logistic regression model with resolved vs unresolved as the outcome and victim age, sex, and race as predictors. 

```{r}
Baltimore_log_reg = 
  US_hom_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, 
    data = .,
    family = binomial())
```

Using the saved object of the `glm` output, we will tidy and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

**OR and its Confidence Interval:**

```{r}
Baltimore_log_reg %>% 
  broom::tidy() %>% 
  filter(term == "victim_raceNon-White") %>% 
  mutate(est_OR = exp(estimate), 
        OR_CI_low = exp(estimate - 1.96*std.error),
        OR_CI_high = exp(estimate + 1.96*std.error)) %>% 
  select(est_OR, OR_CI_low, OR_CI_high) %>% 
  knitr::kable(digits = 3)
```

Now we will run the `glm` function and extract the adjusted odds ratio and confidence interval for each city.

```{r}
city_log_reg = 
  US_hom_data %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(models = map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
         models = map(models, broom::tidy)) %>% 
  select(-data) %>% 
  unnest()
```

```{r}
city_OR_df = city_log_reg %>% 
  filter(term == "victim_raceNon-White") %>% 
  mutate(est_OR = exp(estimate), 
        OR_CI_low = exp(estimate - 1.96*std.error),
        OR_CI_high = exp(estimate + 1.96*std.error)) %>% 
  select(city_state, est_OR, OR_CI_low, OR_CI_high)
```

With this dataframe containing ORs and confidence intervals for each city, we can make a plot.

```{r}
city_OR_df %>% 
  mutate(city_state = forcats::fct_reorder(city_state, est_OR)) %>% 
  ggplot(aes(x = city_state, y = est_OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR_CI_low, ymax = OR_CI_high)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) +
  labs(
    title = "Odds ratio for resolved homicide case comparing non-white to white victims",
    x = "U.S. City",
    y = "Odds Ratio (w/ Confidence Interval)",
    caption = "Data from The Washington Post"
  ) 
```

From the plot we see that most of the odds ratio estimates are below 1; this means that the odds of having a resolved homicide case for a non-white victim are generally lower than the odds of having a resolved homicide case for a white victim. However, a number of confidence intervals (a little less than half) cross 1, making the association between victim race and case resolution less clear in some cities. 

## Problem 2 

First we need to load and clean the data for regression.

Data cleaning steps include making certain variables into factor variables and recoding, in addition to removing the `pnumlbw` and `pnumsga` variable columns since seemingly no data was collected (all values are 0).

```{r}
birthweight_data = read_csv("./data/birthweight.csv")

bwt_data_clean =
  birthweight_data %>% 
  mutate(babysex = as.factor(babysex), 
         babysex = recode_factor(babysex, `1` = "male", `2` = "female"),
         frace = as.factor(frace),
         frace = recode_factor(frace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", `8` = "Other", `9` = "Unknown"),
         malform = as.factor(malform), 
         malform = recode_factor(malform, `0` = "absent", `1` = "present"),
         mrace = as.factor(mrace), 
         mrace = recode_factor(mrace, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Puerto Rican", `8` = "Other")) %>% 
  select(-pnumlbw, -pnumsga)
```

Next a regression model will be constructed.

```{r}
bwt_model = lm(bwt ~ gaweeks + blength + ppwt + smoken + gaweeks*bhead, data = bwt_data_clean)
```

This model is based on the hypothesis that birthweight should be a product of growth indicators such as gestational age and length at birth, as well as "environmental" factors in utero such as mother's pre-pregnancy weight and average number of cigarettes smoked per day during pregnancy. In addition, since gestational age and length at birth are likely to be related, an interaction term between the two is a valid addition to the model. 

**Plot of residuals against fitted values:**

```{r}
bwt_data_clean %>% 
  modelr::add_predictions(bwt_model) %>% 
  modelr::add_residuals(bwt_model) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point(alpha = 0.3) +
  geom_smooth(aes(y = 0), linetype = "dashed") +
  labs(
    title = "Plot of residuals against fitted values",
    x = "Fitted Value",
    y = "Residual",
    caption = "Data from birthweight data of roughly 4000 children"
  ) 
```

Next we compare the above model (`bwt_model`) to two other models (`model one` and `model two`).

```{r}
model_one = lm(bwt ~ gaweeks + blength, data = bwt_data_clean)
model_two = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bwt_data_clean)
```

We will make this comparison in terms of the cross-validated prediction error. 

```{r}
cv_df = crossv_mc(bwt_data_clean, 100)

cv_df =
  cv_df %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(bwt_model = map(train, ~lm(bwt ~ gaweeks + blength + ppwt + smoken + gaweeks*bhead, data = .x)),
         model_one = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
         model_two = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = .x))) %>% 
  mutate(rmse_bwt_mod = map2_dbl(bwt_model, test, ~rmse(model = .x, data = .y)),
         rmse_mod_one = map2_dbl(model_one, test, ~rmse(model = .x, data = .y)),
         rmse_mod_two = map2_dbl(model_two, test, ~rmse(model = .x, data = .y)))
```

**Plot showing distribution of RMSE values for each model:**

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From the plot showing the distribution of RMSE values for each model, we see that mod_one (corresponding to `model_one`, the simplest model that only contains main effects of gestational age and length at birth on birthweight) has the highest RMSE distribution, and is therefore the most inappropriate of the three models. The other models have similar RMSE distributions, but my model (bwt_mod corresponding to `bwt_model`) has more of its bulk slightly lower in the plot, making it a slightly better choice as a regression model.

Therefore, I would ultimately choose the model that incorporates effects of gestational age, length at birth, mother's pre-pregnancy weight, average number of cigarettes smoked per day during pregnancy, and the interaction between gestational age and length at birth on birthweight.





